Demo arXiv template

### 2022-10-08 paper/paper.qmd
build up on what is/ not an index

<!-- Index construction is a way to summarise complicated information (used in environmental data). **it basically says nothing** -->
The complexity of these information may involve a spatial distribution, a temporal scale that defines the frequency of the data, and a multivariate perspective that collects different climate variables.

@@

Various methods are proposed to extract multivariate information on the spatial coverage, across time.
Each individual index follows its own data pipeline and it can be difficult to evaluate how an index can be affected by tweaking parameters in a certain step, rearranging the order of steps, using to different method.

Indices with readily available code to implement have more potential to be computed by researchers, i.e. the package `SPEI` [@SPEI] has implemented the SPI and SPEI index for calculating drought indices. 
[Another package in a different domain].  
However, even if most of the indices have available existing implementation available to users, there would be hundreds of them and it will be burdensome to remember, classify, and compare between.
There is a need for a framework to generally describe how indices can be constructed from spatio-temporal data.

This paper proposes such data pipeline for constructing indices.
The steps involved in the pipeline are general in nature and flexible to be adopted to most index construction for environmental data.

@@

**review indices, climate, drought, economic, UN and other indices**

@@

**who would benefit from this paper: 1) people proposing new indices; 2) people analysing index variation, compare between indices, 3) those making decisions from indices, 4) other statistician, how to modularise steps for anlaysis in your domain.**

@@

standardising
close to variable transformation step

@@

@@

>>>>

<!-- Index construction is a way to summarise complicated information (used in environmental data). **it basically says nothing** -->
<!-- The complexity of these information may involve a spatial distribution, a temporal scale that defines the frequency of the data, and a multivariate perspective that collects different climate variables. -->

@@

<!-- Various methods are proposed to extract multivariate information on the spatial coverage, across time. -->
<!-- Each individual index follows its own data pipeline and it can be difficult to evaluate how an index can be affected by tweaking parameters in a certain step, rearranging the order of steps, using to different method.  -->

<!-- Indices with readily available code to implement have more potential to be computed by researchers, i.e. the package `SPEI` [@SPEI] has implemented the SPI and SPEI index for calculating drought indices.  -->
<!-- [Another package in a different domain].   -->
<!-- However, even if most of the indices have available existing implementation available to users, there would be hundreds of them and it will be burdensome to remember, classify, and compare between. -->
<!-- There is a need for a framework to generally describe how indices can be constructed from spatio-temporal data. -->

<!-- This paper proposes such data pipeline for constructing indices. -->
<!-- The steps involved in the pipeline are general in nature and flexible to be adopted to most index construction for environmental data. -->

@@

  - decision making: need to consider multi-facet for allocating resources, [...]
  - to quantify concept without direct measure. 1) almost impossible to include all the items in the market, indices created from a basket of representative items are useful to characterise the market condition; 2) the concept of interest can't be directly measured, i.e. livability, human developement etc

Integrate multiple information to produce a single number summary (across time).

**Define what is an index, what is not**

Indices quantify a concept of interest that is hard to measure using [...]. 
In finance and economic, the behavior of stock market, indices like Dow Jones Industrial Average, S&P 500, and Nasdaq Composite uses a representative set of companies.
Natural hazard are also indices difficult to define because ....
Various drought, flood indices have been developed to characterise the severity and monitor from the perspective of meteorology, agricultural, hydrology, and also socio-economy. 
Examples of indices are ubiquitous: economic indices for quantifying productibility (GDP?), inflation (CPI), [...]; Social development indices like Human Development Index and Global Liveability Index; diversity indices to characterise the richness, evenness and dominance in ecology

Despite many measurements are called indices, they are actually not indices. 
A direct rate or transformation on a measured variable: birth & death rate computes the total number of new-borns and death per 1000 people. 
Many remote sensing measures are not indices, i.e. Normalized Difference Vegetation Index (NDVI), despite being named as an index, is a mere statistical transformation: the difference of near-infrared (NIR) and red visible channel over its sum.

  <!-- - what is not:  -->
  <!--   - variable transformation in remote sensing, i.e. NDVI is a variable transformation of near-infrared region (NIR) and red visible band.  -->
  <!--   - Another set of measures that are not indices (climate index): Precipitation of driest month, Precipitation of warmest month. These are features -->

- need think through: 
  - is cash rate/ interest rate indices? employment rate? eocnomic indices?

<!-- **review indices, climate, drought, economic, UN and other indices** -->

<!--   - financial and economics index: stock market index (Dow, S&P 500, and Nasdaq), inflation index (CPI, PCE, PPI)  -->
<!--   - indices in newly emerging policy areas such as industrial competitiveness, sustainable development, globalization and innovation. -->
<!--   - Social development indices: Human Development Index (HDI): life expectancy + education + gini -->

@@

  - most indices are presented as calculated numerical or categorical values without open source code accompanied.

  - This causes the problem that if analysts are not agreed with a certain step in the index construction, there is no room for experiment with other possibility.

**who would benefit from this paper**

  - domain scientists who propose new indices and [use multiple indices to compare]
  - decision makers, towards understand the 
  - statisticians, so as to [draw focus/ advocate] on statistical workflow

@@

A specific transformation on the scale of the data

z-score standardising, min-max standardisation into [0, 1] or [0, 100], percentage change on the baseline
close to variable transformation step

@@

Also include weighting

@@

The purpose of normalising is for cross-comparison.
This step can get criticism from analysts for ...

### 2022-10-08 paper/paper.tex
build up on what is/ not an index

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, boxrule=0pt, sharp corners, breakable]}{\end{tcolorbox}}\fi

@@

The complexity of these information may involve a spatial distribution,
a temporal scale that defines the frequency of the data, and a
multivariate perspective that collects different climate variables.

Various methods are proposed to extract multivariate information on the
spatial coverage, across time. Each individual index follows its own
data pipeline and it can be difficult to evaluate how an index can be
affected by tweaking parameters in a certain step, rearranging the order
of steps, using to different method.

Indices with readily available code to implement have more potential to
be computed by researchers, i.e.~the package \texttt{SPEI} (Beguería and
Vicente-Serrano 2017) has implemented the SPI and SPEI index for
calculating drought indices. {[}Another package in a different
domain{]}.\\
However, even if most of the indices have available existing
implementation available to users, there would be hundreds of them and
it will be burdensome to remember, classify, and compare between. There
is a need for a framework to generally describe how indices can be
constructed from spatio-temporal data.

This paper proposes such data pipeline for constructing indices. The
steps involved in the pipeline are general in nature and flexible to be
adopted to most index construction for environmental data.

@@

\textbf{review indices, climate, drought, economic, UN and other
indices}

@@

\textbf{who would benefit from this paper: 1) people proposing new
indices; 2) people analysing index variation, compare between indices,
3) those making decisions from indices, 4) other statistician, how to
modularise steps for anlaysis in your domain.}

@@

standardising close to variable transformation step

@@

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-SPEI}{}}%
Beguería, Santiago, and Sergio M. Vicente-Serrano. 2017. \emph{SPEI:
Calculation of the Standardised Precipitation-Evapotranspiration Index}.
\url{https://CRAN.R-project.org/package=SPEI}.

\leavevmode\vadjust pre{\hypertarget{ref-buja_elements_1988}{}}%
Buja, A, D Asimov, C Hurley, and JA McDonald. 1988. {``Elements of a
Viewing Pipeline for Data Analysis.''} In \emph{Dynamic Graphics for
Statistics}, 277--308. Wadsworth, Belmont.

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, frame hidden, interior hidden, boxrule=0pt, sharp corners, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

@@

\begin{itemize}
\tightlist
\item
  decision making: need to consider multi-facet for allocating
  resources, {[}\ldots{]}
\item
  to quantify concept without direct measure. 1) almost impossible to
  include all the items in the market, indices created from a basket of
  representative items are useful to characterise the market condition;
  2) the concept of interest can't be directly measured,
  i.e.~livability, human developement etc
\end{itemize}

Integrate multiple information to produce a single number summary
(across time).

\textbf{Define what is an index, what is not}

Indices quantify a concept of interest that is hard to measure using
{[}\ldots{]}. In finance and economic, the behavior of stock market,
indices like Dow Jones Industrial Average, S\&P 500, and Nasdaq
Composite uses a representative set of companies. Natural hazard are
also indices difficult to define because \ldots. Various drought, flood
indices have been developed to characterise the severity and monitor
from the perspective of meteorology, agricultural, hydrology, and also
socio-economy. Examples of indices are ubiquitous: economic indices for
quantifying productibility (GDP?), inflation (CPI), {[}\ldots{]}; Social
development indices like Human Development Index and Global Liveability
Index; diversity indices to characterise the richness, evenness and
dominance in ecology

Despite many measurements are called indices, they are actually not
indices. A direct rate or transformation on a measured variable: birth
\& death rate computes the total number of new-borns and death per 1000
people. Many remote sensing measures are not indices, i.e.~Normalized
Difference Vegetation Index (NDVI), despite being named as an index, is
a mere statistical transformation: the difference of near-infrared (NIR)
and red visible channel over its sum.

\begin{itemize}
\tightlist
\item
  need think through:

  \begin{itemize}
  \tightlist
  \item
    is cash rate/ interest rate indices? employment rate? eocnomic
    indices?
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item
  most indices are presented as calculated numerical or categorical
  values without open source code accompanied.
\item
  This causes the problem that if analysts are not agreed with a certain
  step in the index construction, there is no room for experiment with
  other possibility.
\end{itemize}

\textbf{who would benefit from this paper}

\begin{itemize}
\tightlist
\item
  domain scientists who propose new indices and {[}use multiple indices
  to compare{]}
\item
  decision makers, towards understand the
\item
  statisticians, so as to {[}draw focus/ advocate{]} on statistical
  workflow
\end{itemize}

@@

A specific transformation on the scale of the data

z-score standardising, min-max standardisation into {[}0, 1{]} or {[}0,
100{]}, percentage change on the baseline close to variable
transformation step

@@

The purpose of normalising is for cross-comparison. This step can get
criticism from analysts for \ldots{}

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-buja_elements_1988}{}}%
Buja, A, D Asimov, C Hurley, and JA McDonald. 1988. {``Elements of a
Viewing Pipeline for Data Analysis.''} In \emph{Dynamic Graphics for
Statistics}, 277--308. Wadsworth, Belmont.
